---
title: "Correlations"
author: "Francisko de Moraes Rezende"
date: "3/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Running correlations for Gui

## Setup

### Loading packages

```{r loading-pkgs}
list_of_packages <- c("tidyverse", "readxl", "here", "correlation")
new_packages <- list_of_packages[!(list_of_packages %in% installed.packages()[,"Package"])]
if(length(new_packages)) install.packages(new_packages)

library(tidyverse)
library(readxl)
library(here)
library(correlation)
```

### Loading data

```{r loading-data}
data_path <- 
  here::here("data")


files <- 
  dir(path = data_path,
      pattern = "*.xls|*.xlsx")

adsorbed_particles <-
files %>% 
  purrr::map(~ readxl::read_excel(file.path(data_path, .))) %>%
  purrr::set_names(nm = c("particle_count",
                          "macro",
                          "micro",
                          "wetability",
                          "celular_perimeter",
                          "ui"
                          ))
```

### Standardising the sheets

```{r}
adsorbed_particles %>% 
  purrr::pluck("particle_count") %>% 
  dplyr::rename(sp = Sp,
                face = Face,
                frag = Frag,
                site = Site) %>% 
  dplyr::mutate(
    site = dplyr::case_when(
    site == "BR" ~ 1,
    site == "CA" ~ 2,
    site == "VZ" ~ 3,
    site == "CN" ~ 4,
    site == "REF" ~ 5),
    `<2.5` = as.numeric(`<2.5`),
    `2.5-10` = as.numeric(`2.5-10`),
    `10-100` = as.numeric(`10-100`),
    total = `<2.5` + `2.5-10` + `10-100`) -> adsorbed_particles$particle_count
```

```{r}
adsorbed_particles %>%
  purrr::pluck("particle_count") %>%
  dplyr::group_by(site, sp, Rep, face) %>%
  dplyr::summarise(
    `<2.5` = mean(`<2.5`),
    `2.5-10` = mean(`2.5-10`),
    `10-100` = mean(`10-100`),
    total = mean(total)
  ) %>%
  dplyr::ungroup() -> adsorbed_particles$particle_count_micro
```

```{r}
adsorbed_particles %>% 
  purrr::pluck("micro") %>% 
  dplyr::rename(ind = rep) -> adsorbed_particles$micro
```

```{r}
adsorbed_particles %>% 
  purrr::pluck("macro") %>% 
  dplyr::select(-Rq) -> adsorbed_particles$macro
```

```{r}
adsorbed_particles %>% 
  purrr::pluck("micro") %>% 
  dplyr::select(-Sq) -> adsorbed_particles$micro
```


### Combining sheets

#### Combining microroughness sheets

```{r} 
adsorbed_particles %>% 
  purrr::keep(stringr::str_detect(names(.), pattern = "micro")) %>% 
  purrr::reduce(left_join) %>% 
  dplyr::rename(rep = Rep) -> micro_combined
```

#### Combining all vars but microroughness

```{r}
adsorbed_particles %>% 
  purrr::discard(stringr::str_detect(names(.), pattern = "micro")) %>% 
  purrr::reduce(left_join) -> all_but_micro_combined
```
## Analyses

### Defining functions I used

Since there's load of repetition in these analyses, I defined functions to make maintaining the code easier.

The folowing function calculates the pearson correlation between `leaf_char` and `part_size` for a given dataset (`dat`).

```{r}
calculates_correlation <- function(dat, leaf_char, part_size) {

  common_vars <- c("sp", "site", "face", "rep")
  
  dat %>% 
    select(all_of(common_vars), {{leaf_char}}, {{part_size}}) %>% 
    group_by(site, sp, face, rep)  %>%
    summarise(mean({{leaf_char}}), mean({{part_size}})) %>% 
    ungroup() %>% 
    correlation()
}
```


### Microroughness

#### Correcting microroughness sheet

I forgot to change `sp`, `site`, `face` and `rep` to factors so I'll do it now:

```{r}
micro_combined %>% 
  mutate(sp = as.factor(sp),
         site = as.factor(site),
         face = as.factor(face),
         rep = as.factor(rep)) -> micro_combined
```


#### Correlation

```{r}
calculates_correlation(micro_combined, Sa, `<2.5`)
```
